The Elements of Theses

### 2014-09-01 thesis/Background.tex
remove invalide chars

As a result, data deduplication is necessary in order to reduce the storage and bandwidth
costs of backup in a cloud cluster. But adding deduplication at a cloud scale brings new
challenges against previous developed techniques:
\begin{itemize}
\item \textbf{Limited resources.} Without dedicated hardware support in backup appliances, 
VM backup and data deduplication in cloud must not race for CPU and memory resources with
existing VM activities. Compare to dedicated backup appliances which are typically equipped with
tens of GBs of memory and  multiple processors, backup service in the cloud are only allowed to use
no more than a few hundred MBs of memory, and its CPU usage must always stay low.
\item \textbf{Garbage collection for shared data.} Data management for deduplicated storage
is complicated due to the need of tracking use of every shared data block. In order to reclaim
disk space, the reference counts of
shared data blocks will need to be updated whenever a snapshot is added or deleted.
Managing reference counting for data blocks at KB level in a PB level storage is extreme difficult.
\item \textbf{Little semantic information.} For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations take place at the virtual device driver level, which
means no fine-grained file system metadata can be used to determine the changed data. 
\end{itemize}

@@

To that end, this dissertation proposes low-cost deduplication storage architecture that is
collocated with other cloud services. We have learned a lot from 
% The network storage architecture designed fifteen years ago for physical
% servers 
% Backup systems have been developed to use content fingerprints to identify duplicate
% content~\cite{venti02,Rhea2008}.  
% Without an enterprise-class
% alternative, organizations are forced to use traditional solutions, which cannot
% keep up with virtual machines that are dynamic, grow rapidly in number and
% continue to demand new levels of performance and capacity.
various deduplication systems built in the past for different backup scenarios. 
In this chapter we will review the literature and discuss several system design options at high-level.
Section~\ref{back:dedup} introduces the origin and applications of data deduplication.
Section~\ref{back:optm} talks the optimization and tradeoff in designing a deduplication storage system.
Section~\ref{back:arch} discusses design options of backup storage architecture.

\section{Deduplication}
\label{back:dedup}
Backup storage for VM snapshots in the cloud contains PBs of data from many VMs,
the data redundancy in such a data set is very high\cite{Jayaram2011, agrawal07}.
If we treat each snapshot as a single data object, the similarities between snapshots
can not be utilized. Therefore we need better method to break files into data blocks.
In the recent decade, data deduplication techniques are emerged to solve this data
redundancy problem in backup systems.
Most of data de-duplication methods can be categorized by their granularities:

@@

Content-based chunking is first proposed by Manber\cite{similar94}.
He uses Rabin's fingerprints\cite{rabin81, identify00} to sample
data in order to find similar files. This technique computes
fingerprints of all possible substrings of a certain
length in a file and chooses a subset of these fingerprints
based on their values,  e.g., if $(fingerprint(s)\ mod\ a) = b$, where
$s$ is a substring,
$a$ and $b$ are pre-defined values, then the position of substring $s$
is considered as a breakpoint. 
These breakpoints provide 
modification-resistant boundaries that separate the file
into many small chunks. A sequence of hash values of such chunks
is a compact representation of a file, that is then used to compare
against other fingerprinted files.

There are pathological cases in finding breakpoints, for example,
a long extent of zeros may never contain a break point. Thus, LBFS\cite{lbfs01}
introduces min and max thresholds to the chunk size: when scanning the file
sequentially for breakpoints, if the distance between current substring 
and previous breakpoint is less than the minimum threshold, then it is skipped;
if the distance reaches the maximum threshold, then a breakpoint
is enforced. This method also reduces the computation because less substrings
are checked. In order to increase the chance of finding a breakpoint between
two thresholds, Kave\cite{frame05}
proposes the two divisors idea, in addition to have a standard breaking condition,
a weaker breaking condition is used to find backup breakpoints.

@@

\section{Optimization and Tradeoff}
\label{back:optm}
Many previous studies on data deduplication focus on the disk bottleneck problem,
which refers to the key performance challenge in finding duplicate blocks. 
Let's consider a single machine in cluster which hosts many VMs and has 10 TB of virtual disks in total,
if it needs to be backed up daily, then the performance target would be 115 MB/sec on each machine.
Given a block size of 4 KB, a deduplication system must process approximately 28,900 blocks per second.
An in-memory index of all block fingerprints could easily achieve this performance, but the size of the index would limit system size and increase system cost. 
If we use 20 bytes as block fingerprint size, then supporting 10 TB worth of unique blocks, would require 50 GB just to store the fingerprints.
An alternative approach is to maintain an on-disk index of block fingerprints and use a cache to accelerate block index accesses. Unfortunately, a traditional cache would not be effective for this workload. Since fingerprint values are random, there is no spatial locality in the block index accesses. 
With low cache hit ratios, most index lookups require disk operations. If each index lookup requires a disk access which may take 10 msec and 8 disks are used for index lookups in parallel, the write throughput will be about 6.4MB/sec, which is far from the throughput goal of deduplicating at 115 MB/sec for our cloud backup scenario, and don't forget that this goal
is conservative since VM users generally want the backup task to finish in tens of minutes without affecting their normal
application activities.

@@

\begin{itemize}
\item \textbf{Reduce disk index access.} The data domain method ~\cite{bottleneck08} 
uses an in-memory Bloom filter and a locality-based prefetching cache to intercept fingerprint lookup which may
access the on-disk index. Bloom filter\cite{Bonomi2006} helps to tell if a fingerprint is new, and prefetching cache
helps to quickly find consequential duplicates. Improvements to this work with parallelization can be found
 in ~\cite{MAD210,DEBAR,nath08}.
\item \textbf{Index sampling.} A sampled index can be 
put into memory\cite{sparseindex09} or NAND flash\cite{Guo2011} to accelerate fingerprint lookup. 
This sampled index is 100 times or more smaller than the full index so that 
they can be accessed quickly and easily.
Once there's a hit in the sampled index, system loads the fingerprints that are next to the hitted one on disk,
so that it can efficiently catches duplicates based on spatial locality.
\item \textbf{Similarity based sharding.} Several similarity based techniques\cite{shingling97} such like 
Extremely Binning\cite{extreme_binning09, Dong2011, Srinivasan2012, chord01} group big files by similarity metrics
such that similar files are likely to fall into the same group. Then block level deduplication is performed
within the group. While all the group indices are still on the disk, it can deduplicate an entire file
by only load one group index into memory. This approach can easily scale-out, but it misses small amount of duplicates 
because the similarity algorithm does not guarantee always mapping a file to the most similar group.
\end{itemize}

@@

This dissertation studies two low-cost deduplication design options with different tradeoff. 
Chapter~\ref{chap:offline} gives a synchronous parallel deduplication design that
sacrifices the job turnaround time to achieve high throughput and deduplication efficiency.
Chapter~\ref{chap:inline} describes an multi-level selective deduplication solution
that comprises deduplication efficiency to improve throughput and job turnaround time.

\section{Architecture Options} 
\label{back:arch}
The way that backup storage architecture envolves has been following the pace of
primary storage, so we will give a glimpse on primary storage's evolution first.
Organizations start with ﻿building ﻿their ﻿virtualization ﻿infrastructure ﻿using ﻿the﻿
traditional ﻿servers-connected-to-storage-over-a-network ﻿architecture, ﻿which﻿
can’t ﻿adapt ﻿to ﻿the ﻿ever-changing ﻿demands ﻿of ﻿virtualization. ﻿In ﻿addition ﻿to ﻿slow﻿
performance,﻿ network ﻿storage ﻿has ﻿become﻿ the﻿ single﻿ biggest﻿ source ﻿of ﻿cost ﻿and﻿
complexity ﻿in ﻿virtualized ﻿environments.﻿ The ﻿network﻿ storage-based ﻿architecture﻿
worked﻿ well ﻿for ﻿physical﻿ servers ﻿that ﻿served﻿ relatively﻿ static ﻿workloads.﻿
Virtualization,﻿ and﻿ now﻿ Cloud﻿ Computing,﻿ has﻿ made﻿ data﻿ centers﻿ extremely﻿
dynamic\cite{berkeleycloud09};﻿ virtual﻿ machines﻿ are﻿ created﻿ on﻿ the﻿ fly,﻿ move﻿ from﻿ server﻿ to﻿ server﻿ and﻿﻿
depend﻿ heavily﻿ on﻿ shared﻿ resources.﻿ These ﻿characteristics ﻿make ﻿the ﻿management﻿
of ﻿virtual ﻿machines ﻿and ﻿their﻿ underlying ﻿physical﻿ infrastructure ﻿extremely﻿
complex: Data ﻿volumes ﻿are﻿ growing﻿ at ﻿a ﻿rapid ﻿pace﻿ in﻿ the ﻿data ﻿center, ﻿thanks﻿ to ﻿the ﻿ease﻿
of﻿ creating ﻿new﻿ VMs.﻿ In ﻿the﻿ enterprise,﻿ new﻿ initiatives ﻿like ﻿desktop﻿ virtualization﻿
contribute﻿ to ﻿this﻿ trend.﻿ Service﻿ providers﻿ deal﻿ with﻿ an﻿ even﻿ larger﻿ number﻿ of﻿
VMs﻿ as﻿ they﻿ build﻿ data﻿ centers﻿ to﻿ serve﻿ customers﻿ who﻿ can’t﻿ afford﻿ the﻿ cost﻿
and﻿ management﻿ overhead﻿ that﻿ virtualization ﻿requires. ﻿This﻿ growing ﻿pool ﻿of﻿
VMs﻿ is ﻿exerting ﻿tremendous ﻿cost, ﻿performance﻿ and ﻿manageability ﻿pressure﻿ on﻿
the ﻿traditional ﻿architecture ﻿that ﻿connects﻿ compute﻿ to ﻿storage﻿ over ﻿a ﻿multi-hop﻿
network.﻿

Google﻿\cite{googlefs03} and﻿ other﻿ leading ﻿cloud-generation﻿ companies ﻿such﻿ as ﻿Amazon,﻿ Yahoo﻿
and﻿ Microsoft﻿(Azure)﻿\cite{azure} realized ﻿that ﻿a ﻿network-storage﻿ based﻿ approach﻿ would﻿
not ﻿work﻿ for﻿ their﻿ data ﻿centers.﻿ They﻿ built﻿ software﻿ technology﻿ (such ﻿as ﻿Google﻿
File ﻿System)﻿ that﻿ could ﻿glue﻿ a﻿ large﻿ number﻿ of ﻿commodity﻿ servers﻿ with﻿ local﻿
storage ﻿into ﻿a ﻿single﻿ cluster.﻿ This﻿ approach﻿ allowed﻿ Google﻿ to ﻿build ﻿a ﻿converged﻿
compute﻿ and﻿ storage ﻿infrastructure﻿ that﻿ used﻿ commodity﻿ servers﻿ with﻿ local﻿
storage﻿ as ﻿its ﻿building ﻿block\cite{deconstructing05}.﻿ Google﻿ File﻿ System﻿ runs﻿ across﻿ a ﻿cluster ﻿of ﻿servers﻿
and﻿ creates ﻿a ﻿single ﻿pool﻿ of﻿ local﻿ storage﻿ that﻿ can ﻿be﻿ seamlessly﻿ accessed﻿ by﻿
applications﻿ running﻿ on ﻿any ﻿server﻿ in﻿ the ﻿cluster.﻿ It ﻿provides﻿ high﻿ availability﻿
to ﻿applications ﻿by ﻿masking﻿ failures ﻿of ﻿hard﻿ disks ﻿and ﻿even﻿ complete ﻿servers.﻿
Google﻿ File﻿ System ﻿allowed ﻿Google﻿ to﻿ build﻿ data﻿ centers ﻿with﻿ massively ﻿scalable﻿
compute﻿ and﻿ storage,﻿ without ﻿incurring ﻿the ﻿costs﻿ and﻿ performance﻿ limitations﻿
associated ﻿with﻿ network ﻿storage.﻿

>>>>

As a result, data deduplication is necessary in order to reduce the storage and bandwidth
costs of backup in a cloud cluster. But adding deduplication at a cloud scale brings new
challenges against previous developed techniques:
\begin{itemize}
\item \textbf{Limited resources.} Without dedicated hardware support in backup appliances, 
VM backup and data deduplication in cloud must not race for CPU and memory resources with
existing VM activities. Compare to dedicated backup appliances which are typically equipped with
tens of GB of memory and  multiple processors, backup service in the cloud are only allowed to use
no more than a few hundred MB of memory, and its CPU usage must always stay low.
\item \textbf{Garbage collection for shared data.} Data management for deduplicated storage
is complicated due to the need of tracking use of every shared data block. In order to reclaim
disk space, the reference counts of
shared data blocks will need to be updated whenever a snapshot is added or deleted.
Managing reference counting for data blocks at KB level in a PB level storage is extreme difficult.
\item \textbf{Little semantic information.} For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations take place at the virtual device driver level, which
means no fine-grained file system metadata can be used to determine the changed data. 
\end{itemize}

@@

To that end, this dissertation proposes low-cost deduplication storage architecture that is
collocated with other cloud services. We have learned a lot from 
% The network storage architecture designed fifteen years ago for physical
% servers 
% Backup systems have been developed to use content fingerprints to identify duplicate
% content~\cite{venti02,Rhea2008}.  
% Without an enterprise-class
% alternative, organizations are forced to use traditional solutions, which cannot
% keep up with virtual machines that are dynamic, grow rapidly in number and
% continue to demand new levels of performance and capacity.
various deduplication systems built in the past for different backup scenarios. 
In this chapter we will review the literature and discuss several system design options at high-level.
Section~\ref{back:dedup} introduces the origin and applications of data deduplication.
Section~\ref{back:optm} talks the optimization and trade-off in designing a deduplication storage system.
Section~\ref{back:arch} discusses design options of backup storage architecture.

\section{Deduplication}
\label{back:dedup}
Backup storage for VM snapshots in the cloud contains PBs of data from many VMs,
the data redundancy in such a data set is very high\cite{Jayaram2011, agrawal07}.
If we treat each snapshot as a single data object, the similarities between snapshots
can not be utilized. Therefore we need better method to break files into data blocks.
In the recent decade, data deduplication techniques are emerged to solve this data
redundancy problem in backup systems.
Most of data de-duplication methods can be categorized by their granularity:

@@

Content-based chunking is first proposed by Manber\cite{similar94}.
He uses Rabin's fingerprints\cite{rabin81, identify00} to sample
data in order to find similar files. This technique computes
fingerprints of all possible sub-strings of a certain
length in a file and chooses a subset of these fingerprints
based on their values,  e.g., if $(fingerprint(s)\ mod\ a) = b$, where
$s$ is a sub-string,
$a$ and $b$ are pre-defined values, then the position of sub-string $s$
is considered as a breakpoint. 
These breakpoints provide 
modification-resistant boundaries that separate the file
into many small chunks. A sequence of hash values of such chunks
is a compact representation of a file, that is then used to compare
against other fingerprinted files.

There are pathological cases in finding breakpoints, for example,
a long extent of zeros may never contain a break point. Thus, LBFS\cite{lbfs01}
introduces min and max thresholds to the chunk size: when scanning the file
sequentially for breakpoints, if the distance between current sub-string 
and previous breakpoint is less than the minimum threshold, then it is skipped;
if the distance reaches the maximum threshold, then a breakpoint
is enforced. This method also reduces the computation because less sub-strings
are checked. In order to increase the chance of finding a breakpoint between
two thresholds, Kave\cite{frame05}
proposes the two divisors idea, in addition to have a standard breaking condition,
a weaker breaking condition is used to find backup breakpoints.

@@

\section{Optimization and Trade-off}
\label{back:optm}
Many previous studies on data deduplication focus on the disk bottleneck problem,
which refers to the key performance challenge in finding duplicate blocks. 
Let's consider a single machine in cluster which hosts many VMs and has 10 TB of virtual disks in total,
if it needs to be backed up daily, then the performance target would be 115 MB/sec on each machine.
Given a block size of 4 KB, a deduplication system must process approximately 28,900 blocks per second.
An in-memory index of all block fingerprints could easily achieve this performance, but the size of the index would limit system size and increase system cost. 
If we use 20 bytes as block fingerprint size, then supporting 10 TB worth of unique blocks, would require 50 GB just to store the fingerprints.
An alternative approach is to maintain an on-disk index of block fingerprints and use a cache to accelerate block index accesses. Unfortunately, a traditional cache would not be effective for this workload. Since fingerprint values are random, there is no spatial locality in the block index accesses. 
With low cache hit ratios, most index lookups require disk operations. If each index lookup requires a disk access which may take 10 ms and 8 disks are used for index lookups in parallel, the write throughput will be about 6.4MB/sec, which is far from the throughput goal of deduplicating at 115 MB/sec for our cloud backup scenario, and don't forget that this goal
is conservative since VM users generally want the backup task to finish in tens of minutes without affecting their normal
application activities.

@@

\begin{itemize}
\item \textbf{Reduce disk index access.} The data domain method ~\cite{bottleneck08} 
uses an in-memory Bloom filter and a locality-based prefetching cache to intercept fingerprint lookup which may
access the on-disk index. Bloom filter\cite{Bonomi2006} helps to tell if a fingerprint is new, and prefetching cache
helps to quickly find consequential duplicates. Improvements to this work with parallelization can be found
 in ~\cite{MAD210,DEBAR,nath08}.
\item \textbf{Index sampling.} A sampled index can be 
put into memory\cite{sparseindex09} or NAND flash\cite{Guo2011} to accelerate fingerprint lookup. 
This sampled index is 100 times or more smaller than the full index so that 
they can be accessed quickly and easily.
Once there's a hit in the sampled index, system loads the fingerprints that are next to the hit one on disk,
so that it can efficiently catches duplicates based on spatial locality.
\item \textbf{Similarity based sharding.} Several similarity based techniques\cite{shingling97} such like 
Extremely Binning\cite{extreme_binning09, Dong2011, Srinivasan2012, chord01} group big files by similarity metrics
such that similar files are likely to fall into the same group. Then block level deduplication is performed
within the group. While all the group indices are still on the disk, it can deduplicate an entire file
by only load one group index into memory. This approach can easily scale-out, but it misses small amount of duplicates 
because the similarity algorithm does not guarantee always mapping a file to the most similar group.
\end{itemize}

@@

This dissertation studies two low-cost deduplication design options with different trade-off. 
Chapter~\ref{chap:offline} gives a synchronous parallel deduplication design that
sacrifices the job turnaround time to achieve high throughput and deduplication efficiency.
Chapter~\ref{chap:inline} describes an multi-level selective deduplication solution
that comprises deduplication efficiency to improve throughput and job turnaround time.

\section{Architecture Options} 
\label{back:arch}
The way that backup storage architecture envolves has been following the pace of
primary storage, so we will give a glimpse on primary storage's evolution first.
Organizations start with building their virtualization infrastructure using the
traditional servers-connected-to-storage-over-a-network architecture, which
can’t adapt to the ever-changing demands of virtualization. In addition to slow
performance, network storage has become the single biggest source of cost and
complexity in virtualized environments. The network storage-based architecture
worked well for physical servers that served relatively static workloads.
Virtualization, and now Cloud Computing, has made data centers extremely
dynamic\cite{berkeleycloud09}; virtual machines are created on the fly, move from server to server and
depend heavily on shared resources. These characteristics make the management
of virtual machines and their underlying physical infrastructure extremely
complex: Data volumes are growing at a rapid pace in the data center, thanks to the ease
of creating new VMs. In the enterprise, new initiatives like desktop virtualization
contribute to this trend. Service providers deal with an even larger number of
VMs as they build data centers to serve customers who can’t afford the cost
and management overhead that virtualization requires. This growing pool of
VMs is exerting tremendous cost, performance and manageability pressure on
the traditional architecture that connects compute to storage over a multi-hop
network.

Google\cite{googlefs03} and other leading cloud-generation companies such as Amazon, Yahoo
and Microsoft(Azure)\cite{azure} realized that a network-storage based approach would
not work for their data centers. They built software technology (such as Google
File System) that could glue a large number of commodity servers with local
storage into a single cluster. This approach allowed Google to build a converged
compute and storage infrastructure that used commodity servers with local
storage as its building block\cite{deconstructing05}. Google File System runs across a cluster of servers
and creates a single pool of local storage that can be seamlessly accessed by
applications running on any server in the cluster. It provides high availability
to applications by masking failures of hard disks and even complete servers.
Google File System allowed Google to build data centers with massively scalable
compute and storage, without incurring the costs and performance limitations
associated with network storage.
